### 2012-08-01 latex/sdext/src/document.tex
first commit

\documentclass[twocolumn]{article}

% config
\usepackage[brazil]{babel}
%\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{setspace}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textfit}
%\usepackage[numbers, sort]{natbib}
\usepackage[natbib=true,sorting=none, firstinits=true]{biblatex}
\usepackage{url}
\usepackage{color}  
\usepackage{wrapfig}
\usepackage[top=30mm,bottom=20mm,left=30mm,right=20mm,twoside]{geometry}

\usepackage{array}
\usepackage{multirow}

\newcommand\MyBox[1]{
        \fbox{\lower0.75cm
                \vbox to 1.7cm{\vfil
                        \hbox to 1.7cm{\hfil\parbox{0.7cm}{#1}\hfil}
                \vfil}%
        }%
}

\hypersetup{
        colorlinks=true,
        citecolor=black,
        filecolor=black,
        linkcolor=black,
        urlcolor=black
}

\bibliography{references}

\title{}

\author{Francisco Morgani Fatore}

\date{\today}
\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introdução}\label{sec:intro}

\section{Trabalhos Relacionados}\label{sec:related}

\section{Metodologia}\label{sec:metodology}

\section{Resultados}\label{sec:resuls}

\section{Conclusão}\label{sec:conclusion}

\printbibliography

\end{document}